---
title: "R basics"
output: word_document
---

# 1.1 About R
- R came from S
- R works in memory
    + Microsoft has out-of-memory processing capabilities 
    + There's some other potential options
- R is quirky AF
- R is case sensitive
- There's a package for that!

# 1.2 R packages
We can install from CRAN
```{r eval=FALSE}
install.packages("tidyverse")
```

We can make packages available to us using `library()` function.

```{r}
library("tidyverse")
```

```{r}
tidyverse_packages()

```

# 1.3 Vectors
1 dimensional objects

```{r}
colors <- c("red","blue","green")
colors2 = c("orange","purple")

nums <- 1:10
nums

about <- c(firstname="Tove",lastname="Hjelm")
about
```
Extracting values uses a grid system

```{r}
colors[1]
colors[-1]
colors[c(1,3)]
colors[1:2]
colors[-c(1,2)]

```
# 2.1 Tidyverse

To clean and prepare data from different sources.

```{r}
library(tidyverse)

# Change "data" to where your files are.
# Remove the col_types bit if your columns
# are fairly type safe.
list.files("data", full.names = TRUE) %>%
  map_df(read_csv, col_types = cols("c")) %>% 
  bind_rows() %>% 
  nrow()
```

# 2.2 Caret package
Create datasets with training data (features) and training outcome (what we want to predict)
Create datasets to test the model. Test data with features that are used to predict, and test outcome that has the actual cylender of the data we test.
method = what model we are using, neural network, regression etcetera. 
preprocess = olika typer av forbearbetning. pca = vilka variabler som ar viktiga, vilka faktorer som paverkar cylinderantal.

```{r}
library(caret)

## Loading required package: lattice

## 
## Attaching package: 'caret'

## The following object is masked from 'package:purrr':
## 
##     lift
## To predict cylinder (column 2) from other features.


training_data    <- mtcars[1:20,-2]
training_outcome <- mtcars[1:20,2]
test_data        <- mtcars[21:32,-2]
test_outcome     <- mtcars[21:32,2]

model <- train(training_data, training_outcome,
               method="lm",
               preProcess = c("scale","center","pca"))
predictions <- predict(model, test_data)
predictions
```
#2.3 Visualizations

```{r}
library(ggplot2)
library(datasauRus)

myPlot<- ggplot(datasaurus_dozen, aes(x,y)) +
           geom_point()
myPlot
```
```{r}
myPlot + facet_wrap(~dataset)
```
Uses the myPlot we defined but with the simpsons_paradox dataset instead of datasaurus_dozen. Makes it possible to reuse models with other data. 

(Simpson’s Paradox is the phenomenon where high level statistics provide one conclusion but evaluating sub-groups within the data provides a very different conclusion.)

```{r}
myPlot %+% simpsons_paradox
```
# 32. Visualizations, Common charts

```{r}
library(ggplot2)
library(datasauRus)
library(cowplot)

ggplot(datasaurus_dozen, aes(x = x, y = y)) + geom_point()

p <- ggplot(datasaurus_dozen, aes(x = x, y = y, color=dataset)) + geom_line() + facet_wrap(~dataset)
p

ggplot(data = mtcars, aes(x = rownames(mtcars), y = mpg)) + geom_col() + theme(axis.text.x = element_text(angle = 90))

plot1 <- ggplot(data = simpsons_paradox, aes(x = x, y = y)) + geom_point()
plot2 <- ggplot(data = mtcars, aes(x = am)) + geom_bar()
ggplot(data = mtcars, aes(x = rownames(mtcars), y = mpg)) + geom_col()
ggplot(data = simpsons_paradox, aes(x = x, y = y, colour = dataset)) + geom_jitter(width = 5)
ggplot(data = simpsons_paradox, aes(x = x, y = y, colour = dataset)) + geom_point(alpha = 0.5)
ggplot(data = simpsons_paradox, aes(x = x, y = y, size = dataset, colour = dataset, 
    shape = dataset, alpha = dataset)) + geom_point()
ggplot(data = simpsons_paradox, aes(x = x, y = y, group = dataset, colour = dataset)) + 
    geom_line()


plot_grid(plot1,plot2)

```

## Exercises
```{r}

#Iris
ggplot(data = iris, aes(x = Sepal.Width, y = Sepal.Length, color = Species, size = Petal.Width+Petal.Length)) + geom_point(alpha = 0.8)

#Barchart, get max for all y values
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarise(y=max(y)) %>%
  ggplot(aes(x=dataset, y=y)) +
  geom_col()


```
## 33.Visualizations, Density charts

```{r}
ggplot(data = mtcars, aes(x = wt)) + geom_histogram()
ggplot(data = mtcars, aes(x = wt)) + geom_density()
ggplot(data = mtcars, aes(x = as.factor(am), y = wt)) + geom_violin()
ggplot(data = mtcars, aes(x = wt, mpg)) + geom_bin2d()
```

## 34. Faceting
```{r}
ggplot(data = datasaurus_dozen, aes(x = x, y = y)) + geom_point() + facet_wrap(~dataset)

ggplot(data = datasaurus_dozen, aes(x = x, y = y)) + geom_point() + facet_wrap(~dataset, 
    nrow = 5)
```
# Exercises
Create a facet scatterplot for simpsons_paradox

```{r}
library(ggthemes)
ggplot(data = simpsons_paradox, aes(x = x, y = y)) + geom_point() + facet_wrap(~dataset) + theme_classic()
```
## Packages that are good for Data Exploration

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("krupanss/IEDA")
# skimr::skim(dataset) to get a good overview of a dataset
# GGally::ggpair(dataset) to pair all variables with eachother

```
## Clean Ecommerce-data (Description contains a stringvalue with \x, R does not like that)

```{r}
ecommerce %>%
  mutate(Description=str_remove_all(Description,"[^[:alnum:]]")) %>% 
  skimr::skim()

```

```{r}
ecommerce %>% 
  ggplot(aes(x = Quantity)) + geom_density()
```
## Create awesome data explorer data
```{r}
#library(DataExplorer)
#DataExplorer::create_report(dataset, "filename.html")
```



