---
title: "R basics"
output: pdf_document
---

# About R
- R came from S
- R works in memory
    + Microsoft has out-of-memory processing capabilities 
    + There's some other potential options
- R is quirky AF
- R is case sensitive
- There's a package for that!

# R packages
We can install from CRAN
```{r eval=FALSE}
install.packages("tidyverse")
```

We can make packages available to us using `library()` function.

```{r}
library("tidyverse")
```

```{r}
tidyverse_packages()

```

# Vectors
1 dimensional objects

```{r}
colors <- c("red","blue","green")
colors2 = c("orange","purple")

nums <- 1:10
nums

about <- c(firstname="Tove",lastname="Hjelm")
about
```
Extracting values uses a grid system

```{r}
colors[1]
colors[-1]
colors[c(1,3)]
colors[1:2]
colors[-c(1,2)]

```
## Caret package
Create datasets with training data (features) and training outcome (what we want to predict)
Create datasets to test the model. Test data with features that are used to predict, and test outcome that has the actual cylender of the data we test.
method = what model we are using, neural network, regression etcetera. 
preprocess = olika typer av forbearbetning. pca = vilka variabler som ar viktiga, vilka faktorer som paverkar cylinderantal.

```{r}
library(caret)

## Loading required package: lattice

## 
## Attaching package: 'caret'

## The following object is masked from 'package:purrr':
## 
##     lift
## To predict cylinder (column 2) from other features.


training_data    <- mtcars[1:20,-2]
training_outcome <- mtcars[1:20,2]
test_data        <- mtcars[21:32,-2]
test_outcome     <- mtcars[21:32,2]

model <- train(training_data, training_outcome,
               method="lm",
               preProcess = c("scale","center","pca"))
predictions <- predict(model, test_data)
predictions
```

